\LaTeXe\ guide for the Journal of Functional Programming

### 2019-04-25 papers/jfp/7-experience.tex
Work on S7

As a set of authors, we've written a paper about the \Shake build system \cite{mitchell2012shake},
applying that build system to GHC \cite{hadrian}, and a paper about the structure of build
systems \cite{mokhov2018buildsystems}. The original purpose of that final paper was to create
``Cloud Shake'', and use it for the GHC build system. In this section we tie these papers together,
reflecting on our experience after all these papers.

@@

The original design of \Shake hasn't changed since the initial paper. Since that paper there have been roughly 5,000 commits to the \Shake project\footnote{\url{https://github.com/ndmitchell/shake}}. They add concepts like resources (if two rules content from a single external resource), rewrite serialisation to be faster, and documentation including a website, and add lots of tests. The biggest change in that time period was moving from blocking threads to continuations for the suspending scheduler. The most visible change is switching \hs{*>} for \hs{\%>}\footnote{A conflicting \hs{*>} operator was added to the Haskell \hs{Prelude}.}, but almost all external and internal details remain the same.

We consider the lack of change to suggest that \Shake is based on underlying principles -- principles we can now name and describe as a consequence of \cite{mokhov2018buildsystems}.

@@

Converting \Shake into Cloud \Shake wasn't a difficult process armed with a roadmap. The key was the introduction of two new functions:

@@

These functions are suspiciously like \hs{recordCT} and \hs{constructCT} from \S\ref{sec-constructive-traces}, with their differences perhaps the most illustrative of the changes required. (We have made the replacements, like replacing \hs{Key} for \hs{k}, to reduce irrelevant differences.)

@@

\subsection{Experience from building GHC with Cloud \Shake}

\begin{itemize}
\item Absolute paths and system binaries
\item Deferred materialisation. Different invariants for (a) local builds, (b) cloud sharing (you must list all the things you produce), (c) sharing + deferred materialisation (you must declare all the things you consume). Interaction with early cut-off.
\item  Cloud stuff needed tracing infrastructure to expose dependencies.
\end{itemize}

Can reduce times a lot.

Needs to get much better dependencies.

Also been deployed by Standard Chartered.

>>>>

As a set of authors, we've written a paper about the \Shake build system \cite{mitchell2012shake},
applying \Shake to building GHC \cite{hadrian}, and a paper about the categorisation of build
systems \cite{mokhov2018buildsystems}. The original motivation of the research in that final paper was to create
``Cloud Shake'', and use it for the GHC build system. In this section we tie these papers together,
reflecting on our experience since these papers.

@@

The original design of \Shake hasn't changed since the initial paper. Since that paper there have been roughly 5,000 commits to the \Shake project\footnote{\url{https://github.com/ndmitchell/shake}}. They add concepts like resources (if two rules content from a single external resource), rewrite serialisation to be faster, documentation including a website\footnote{\url{https://shakebuild.com}}, and add lots of tests. The biggest change in that time period was moving from blocking threads to continuations for the suspending scheduler. The most visible change is switching \hs{*>} for \hs{\%>}\footnote{A conflicting \hs{*>} operator was added to the Haskell \hs{Prelude}.}, but almost all external and internal details remain the same.

We consider the lack of change to suggest that \Shake is based on fundamental principles -- principles we can now name and describe as a consequence of \cite{mokhov2018buildsystems}.

@@

Converting \Shake into Cloud \Shake wasn't a difficult process once armed with a roadmap. The key was the introduction of two new functions:

@@

These functions are suspiciously like \hs{recordCT} and \hs{constructCT} from \S\ref{sec-constructive-traces}, with their differences perhaps the most illustrative of the changes required. (We have made some replacements from full \Shake, like replacing \hs{Key} for \hs{k}, to reduce irrelevant differences.)

@@

\subsection{Experience from using Cloud \Shake}

While we expected the GHC build system to be the first to take advantage of Cloud \Shake, we were actually beaten to it by Standard Chartered who report\footnote{\url{https://groups.google.com/d/msg/shake-build-system/NbB5kMFS34I/mZ9L4TgkBwAJ}}:

\begin{quote}
Thanks for the symlinks release, we just finished upgrading this build system to use --share. ...
Building from scratch with a warm cache takes around 5 seconds, saving us up to 2 hours. Not bad!
\end{quote}

We have also got the GHC build working with Cloud \Shake, resulting in build times of N minutes instead of M minutes. Due to aspects like \hs{configure} we don't see the dramatic reductions reported by Standard Chartered. Converting to a build suitable for sharing is not overly onerous, but nor is it trivial. Some things that require attention:

\begin{description}
\item[Irrelevant differences] A common problem is that you don't get shared caching when you want it. To get caching, it is important that all machines are identical for the aspects that involve the build system. In particular, that means that all paths must be relative, and all dependencies on system binaries (e.g. \hs{gcc}) should either be controlled by the build system (e.g. using \Nix) or have tolerence to account for minor machine variations.
\item[Insufficient produced files] A build rule must declare which files it produces, so these are stored in the cache and can be restored from the cache. If you don't use a remote cache then it is permissible to have files not known to the build system which change at the same time as files that are known. Once using a remote cache any file that is produced as a side-effect of a rule and then later used \emph{must} be declared. Failure to declare such files means that if some step produces a file, and a later step relies on it, that step breaks. In practice, most issues encountered during the move to Cloud builds for GHC were a lack of declared produced files.
\item[Insufficient dependencies] While insufficient dependencies are often a problem, the move to a Cloud build makes them even more critical, as missing dependencies are likely to mean a result is \emph{never built locally}, rather than just build locally only once. Furthermore, dependencies in a Cloud setting must be more complete. For example, using the build system from Figure \ref{fig-make}, with a local build system without early cutoff, if the rule to produce \hs{main.exe} also made use of \hs{main.c} but didn't declare so, that isn't a problem. With the addition of early cutoff, if \hs{main.c} changes in an important way but \hs{main.o} doesn't update that causes issues, but is a somewhat rare occurance. With Cloud builds that becomes somewhat more common.
\end{description}

To help with the final two issues -- insufficient dependencies and produced files -- we have further enhanced the \Shake lint modes, coupling them to a program called \hs{FSATrace}, which detects which files are read/written by a command line execution. Such information has been very helpful in making the GHC build Cloud ready.
